A few stickleback suffice for the transport of alleles to new lakes

### 2019-07-04 Stickleback_Paper.tex
peter edits

Threespine stickleback populations provide a striking example of local adaptation to divergent habitats while still being connected by recurrent gene flow. 
These small fish occur in marine and freshwater habitats throughout the Northern Hemisphere, and in numerous cases the smaller freshwater populations have been established ``de novo'' from marine colonists. 
Previous results have found similar phenotypes between independently evolved freshwater populations are derived largely from the same standing genetic variants found to be Identical-by-decent.
Geographic isolation preventing direct migration between the freshwater populations strongly suggests that these shared adaptive alleles are transported through the widely distributed marine population.
However it is still largely unknown how gene flow, recombination, and selection jointly impact standing variation to fuel adaptation.
Here we use genealogical simulations to determine the levels of gene flow that best match observed patterns of allele sharing among habitats in stickleback.
We aim to better understand how gene flow and local adaptation in large metapopulations determine the speed of adaptation and re-use of standing genetic variation. 
In our simulations we find that repeated adaptation uses a shared set of alleles that are maintained at low frequency by migration-selection balance in oceanic populations. 
This process occurs over a realistic range of intermediate levels of gene flow that match previous empirical population genomic studies in stickleback. 
Examining these simulations more deeply reveals that lower levels of gene flow leads to slow, independent adaptation to different habitats, whereas higher levels gene flow leads to significant mutation load -- but an increased probability of successful population genomic scans for locally adapted alleles. 
Surprisingly, we find that the genealogical origins of most freshwater adapted alleles can be traced back to the original generation of marine individuals that colonized the lakes, as opposed to subsequent migrants. 
These simulations provide deeper context for existing studies of stickleback evolutionary genomics, and guidance for future empirical studies in this model.
More broadly, our results support existing theory of local adaptation but extend it by more completely documenting the genealogical history of adaptive alleles in a metapopulation.

@@

The canonical model for the genetics of adaptation has long been the sequential fixation of new mutations 
\citep{smith_haigh_1974, Endler_1977, Orr_2005}.
While it has proved to be a useful baseline for understanding the genetic variation we see in species today, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures 
\citep{Lai2152, Schrider2017}. 
In particular, empirical studies have increasingly identified the need to more deeply incorporate standing genetic variation (SGV) into adaptation dynamics for metapopulations inhabiting an array of habitats 
\citep{Hermisson_2008, Barrett2008}. 
Populations experiencing diverse selective pressures while still exhibiting significant gene flow often result in more complex genomic signals that are still not fully understood 
\citep{Charlesworth_1997, Charlesworth_2003, Nosil_2009, Flaxman_2013, Samuk_2017}. 
Concurrently, a growing number of empirical studies have identified instances of convergent evolution using SGV 
\citep{Schrider2017, Barrett2008, Nelson_2018, Nelson_2019, Bassham_2018}. 
However, it is still not clear how variation in evolutionary processes - such as gene flow, recombination, selection and mutation -  can promote the maintenance and re-use of SGV, particularly during colonization and adaptation to new environments
\citep{nelson2017ancient, Pritchard_2010, yeaman2011genetic, Schrider2017}. 
It is similarly unclear whether variation in these evolutionary processes can determine the genetic architecture of evolving traits via SGV. 
While empirical studies in the field and lab are essential, modeling specific systems can help us understand the nature of standing genetic variants to fuel local adaptation as well as the genomic signals we can expect to see.

The ancestral marine form of threespine stickleback fish has given rise to millions of independently derived freshwater populations in recently de-glaciated regions around the Northern Hemisphere 
\citep{Bell_1994, Thompson_1997, Cresko_2007}. 
This model organism has provided some of the earliest data in a recent flood showing the heterogeneous nature of divergence across genomes 
and the much more extensive use of SGV than once thought 
\citep{schluter2009genetics, Roesti_2014, Nelson_2018, Bassham_2018, Nelson_2019, Hohenlohe_2010, Terekhanova_2014, Marques_2016}.  
While geographic isolation often prevents direct migration between freshwater populations, stickleback in them frequently evolve similar phenotypes 
\citep{Cresko_2004, Colosimo_2004}. 
The most recent evolutionary genomic studies on stickleback document that while the overall dynamic of local adaptation to marine and freshwater habitats has been occurring for millions of years 
\citep{Nelson_2018}, independent local adaptation of marine individuals to freshwater environments has been observed to take place in just tens of generations \citep{Terekhanova_2014, Lescak2015evolution, Bassham_2018}. 
For example, in 1964 the Great Alaskan Earthquake caused an uplift of many islands and coastal regions throughout the Gulf of Alaska. 
Studies of stickleback populations on uplifted Middleton Island showed that newly created freshwater ponds were invaded by the surrounding marine population of stickleback which evolved the freshwater syndrome of phenotypes in less than 50 years 
\citep{Lescak2015evolution, Bassham_2018}. 
Amazingly, despite very little change in a majority of the genome, nearly a quarter of it diverged significantly in this same time in a pattern that mirrored previous genomic studies of freshwater stickleback populations that have been geographically separated for thousands of years 
\citep{Hohenlohe_2010, Bassham_2018}.

But how can evolution occur at such a rapid pace?
Waiting for new mutations to arise in each lake or pond would take much longer than the decades since the Alaskan earthquake. 
Even more improbable is having divergence cluster in such similar genomic regions across independent populations. 
An alternative hypothesis, supported by the most recent population genomic analyses, is that the majority of alleles important for freshwater adaptation are maintained in the marine individuals due to recurrent gene flow from freshwater back in to marine populations. 
In 2009 \citet{schluter2009genetics} proposed a conceptual model they termed the ``transporter''- hypothesis to describe the process of alleles beneficial in freshwater environment being maintained at migration-selection balance in the larger oceanic population and thus available for subsequent adaptation to new freshwater habitats.
The first clear example of the global reuse of SGV in stickleback was the gene \textit{eda} which has been shown to be an important regulator for the number of bony lateral plates 
\citep{Colosimo_2004, Cresko_2004, Colosimo_2005}. 
While the low lateral plate version of this gene arose millions of years ago, it is found in much younger freshwater ponds around the Northern Hemisphere 
\citep{O'Brian_2015}. 
More recently, genome-wide haplotype analyses have provided evidence that \textit{most} regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{nelson2017ancient}.

While this growing body of population genomic data on stickleback evolution support the transporter hypothesis, a number of questions remain. 
What are the actual population sizes, migration rates, and fitness differentials consistent with this hypothesis? 
How many differentially selected alleles exist, how many are used at any one time, and how are they arranged within the genome? 
A curious natural history observation underlying many of these questions is the fact that some newly formed freshwater habitats, such as the ponds on Middleton Island, are quite small and presumably the number of initial marine migrants is few.

Here, we use forward simulations implemented in SLiM that incorporate linkage and realistic length genomes to model the stickleback metapopulation and address these questions \citep{haller2017slim,haller2018slim3}. 
We ask how variation in amount of gene flow affects the genetic architecture of local adaptation to newly created freshwater ponds.
Because we are documenting the entire genealogy of all alleles, we can determine the distribution and abundance of haplotypes across all in marine and freshwater populations to know the timing and proportion of potentially adaptive alleles that are actually utilized in each population. 
In addition, we can document how these adaptive alleles are distributed across the genomes, and as a consequence determine how this may affect the efficacy of genome scans for between-habitat differentiation. 
We find that only a few stickleback migrants per lake per generation are sufficient to maintain SGV in the ocean, allowing new freshwater populations to adapt in only tens of generations. 
We also find that continued gene flow of marine individuals into recently established lakes is not necessary for adaptation as the majority of alleles are present in the initial marine colonizers. 
Finally, we find that this rapid, local adaptation can only occur over an intermediate range of levels of gene flow. Low gene flow prevents variation from being transported, while high gene flow prevents local adaptation. 
Together these simulation findings enhance and extend the existing rich empirical work on stickleback evolutionary genomics.

@@

To explore these questions, we used SLiM \citep{haller2017slim,haller2018slim3} to implement complex forward-time simulation of populations of individuals with explicitly represented genomes in which selection acted upon a single continuous quantitative trait. 
The details of the model were motivated by current understanding of threespine stickleback evolutionary history and demography, and included divergent selection in the two habitats which had substantial spatial structure. 
Certain aspects of the model remain simplistic due to computational constraints.
Possibly the most important caveat is that simulated population sizes are much smaller than the census size of the threespine stickleback populations observed in nature (see the Discussion for more on this).
\jgg{I don't know where we should talk about our small population sizes ... Is this even worth mentioning here?}
\wac{I don't think it's necessary because the census population sizes of stickleback actually due vary widely, so we're not that far off. If anything you could say that we focused on smaller populations because of our questions about how many migrants are needed}

\paragraph{Habitat and geography}

Our simulations include two habitat types -- Marine and Freshwater -- defined by the nature of their selective pressures, each with 5,000 diploid individuals. 
The arrangement of these habitats, depicted in Figure~\ref{fig:Geo}, roughly models a set of freshwater habitats along a stretch of coastline. 
The marine habitat is a continuous, one-dimensional range of length 25 units, while the freshwater habitat is divided into 25 discrete subpopulations (which we call ``lakes''), each connected to the marine habitat at regularly spaced intervals (positions $i - 1/2$ for $1 \le i \le 25$).

Divergent selection is mediated by a single quantitative trait with different optima in marine and freshwater habitats. 
This situation roughly models the cumulative effect of the various phenotypes thought to be under divergent selection between the habitats, such as armor morphology, body size, craniofacial variation and opercle shape (among many other traits). 
The optimal trait values in the marine and freshwater habitats are $+10$ and $-10$ respectively, and fitness of a fish with trait value $x_\text{ind}$ in a habitat with optimal value $x_\text{opt}$ is determined by a Gaussian kernel with standard deviation 15, i.e.,

@@

We use SLiM to simulate a Wright--Fisher population with non-overlapping generations and a fixed population size of 5,000 diploid individuals in each habitat. 
Each generation, the two parents of each new offspring are chosen proportional to their fitness (all individuals are hermaphroditic), and the contributing genomes are produced by Poisson recombination with an average of one crossover per chromosome per generation ($10^{-8}$ per locus per generation). 
Since the total population across \emph{all} 25 lakes is fixed at 5000, and the Wright--Fisher model assumes unrealistic global population regulation, we normalize the fitnesses of each individual such that approximately 200 offspring are generated in each lake, each generation. To do this, we divide fitness values of each freshwater individual by the mean fitness in their lake, so that the mean fitnesses of all lakes are equal before selection happens.

@@

We used SLiM's ability to record \emph{tree sequences} \citep{haller2018treesequence,Kelleher2016} to output the genealogical history of all individuals at the time of introduction of new lakes, at the time of adaptation, and at the end of the simulation. This allowed us to directly query the true origins of adaptive alleles. 
In addition, it allowed for much larger simulations by avoiding the computationally expensive task of simulating neutral mutations which were retroactively added to the gene trees at a rate of $10^{-8}$ per locus per generation, as described in \citet{kelleher2018efficient}.
The output tree sequence from each simulation allows us to explore the origin of the genetic basis of adaptation in the new lakes. 
To do this, we constructed the genealogical tree relating all extant chromosomes at each locus along the genome. 
Using these trees we classified each adaptive allele, in each genome in the new lakes at the time of adaptation, into four categories:

@@

We used neutral mutations to calculate $F_{ST}$, calculated on a per-locus basis and then averaged in windows as a measure of between-population relative differentiation. 
Concretely, if $p_f$ and $p_m$ are the frequencies of a given mutant allele in the freshwater and marine habitats, respectively, and $\bar p = (p_f + p_m)/2$, then we compute $F_{ST}$ for that mutation as $1 - p_f p_m / (\bar p (1-\bar p))$.

\section*{Results \& Discussion}

@@

To observe the impact of gene flow on selection in the new lakes, we varied the ocean--lake migration rate, $m$, across separate simulations from $5 \times 10^{-5}$ to $5 \times 10^{-1}$. 
Below, we often refer to these migration rates in terms of the number of migrants per lake, per generation, which we denote $M$. Since each lake contains 200 individuals, $M=200m$. 
Many aspects of adaptation changed substantially across this range, including the speed of adaptation, degree of sharing of adaptive alleles between lakes, and the population genetic signals left behind. 
At very low rates of gene flow, each new lake's population adapted almost completely independently through \emph{de novo} mutation, which took a very long time ($\approx 20,000$ generations). 
At very high rates of gene flow, local adaptation was constrained by the large influx of locally maladaptive alleles \citep{bolnick2007natural}. 
Between these two extremes, genetic variation that allowed adaptation to freshwater habitats could move relatively easily between lakes. 
Perhaps surprisingly, only a few migrants per generation from the lakes to the ocean were needed to maintain sufficient genetic variation in the ocean to dramatically accelerate adaptation in new lakes.

@@

\begin{figure}
    \begin{center}
        \includegraphics{Final_Plots/Pheno_Time.pdf}
          \caption{ 
                Mean individual trait values in the marine habitat (blue line),
                the original lakes (yellow lines; average in orange),
                and the new lakes (light green lines; average in dark green),
                across the course of two simulations, with migration rates of
                \textbf{(top)} $M=0.01$ and
                \textbf{(bottom)} $M=0.1$ 
                migrants per lake per generation, respectively.
                Optimal trait values in the two habitats are at $\pm 10$.
                    Analogous plots for other migration rates
                    are shown in Figure \ref{fig:phenotype_time_supp}.
        }
          \label{fig:phenotype_time}
    \end{center}
\end{figure}

As shown in Figure~\ref{fig:MeanPhenotype}, local adaptation occurred in all simulations with the exception of the highest gene flow at which half of each population was composed of migrants. 
Figures \ref{fig:phenotype_time} and \ref{fig:phenotype_time_supp} show how mean trait values in freshwater and marine populations diverged over time, until the trait means were close to the optimal values in each habitat. 
Establishment of new alleles in the lakes is visible in the first few thousand generations as jumps in the mean trait value which move the trait by an amount of order 1 every few hundred generations (Fig. \ref{fig:phenotype_time}).
At the lowest rate of gene flow, $M = 0.01$, differences at around 20 commonly polymorphic sites (about 10 that shift the trait in each direction) were responsible for most of the adaptive differences between freshwater and marine habitats. 
As expected, increasing migration rate decreased differentiation between habitats: as seen in Figure~\ref{fig:Fst}, $F_{ST}$ between marine and freshwater habitats at neutral sites steadily declines as migration increases.

\begin{figure}
    \begin{center}
        \includegraphics[width=\textwidth]{Final_Plots/Fst_Genome_faa.pdf}
          \caption{
        Average $F_{ST}$ in windows of 500Mb between:
        \textbf{(top)} marine habitat and old lakes;
        \textbf{(middle)} marine habitat and new lakes; and
        \textbf{(bottom)} old and new lakes.
        Each plot shows $F_{ST}$ values for a separate simulation,
        with columns corresponding to increasing gene flow from left to right.
        $F_{ST}$ is calculated between all marine individuals, and all lakes pooled together.
        All locations of pre-existing freshwater adapted alleles have been highlighted by 
        pink dashed vertical lines with an alpha value of 0.3, so darker shades of pink imply more 
        freshwater adapted alleles at that location.
     } \label{fig:Fst}
    \end{center}
\end{figure}

Adaptation occurred much more quickly at higher migration rates, both in the old and new sets of lakes. We measured this ``time to adaptation'' as the number of generations until average trait values in old and new lakes were within 0.5 of each other, shown in Figure~\ref{fig:TimeToAdaptation} for different rates of gene flow. Adaptation of new lakes took over 18,000 generations at the lowest rate of $m = 0.01$, while at $m = 1$ migrant per, lake per generation, new lakes managed to adapt in just under 60 generations.

@@

At low migration rates, the \emph{initial} period of adaptation takes roughly 25 times longer for lakes than it does for the ocean. 
This difference occurs because at low migration rates, each lake must wait for its own novel mutations to arise in order to adapt. 
Because the marine habitat is continuous, with 25 times more individuals than any one lake, there is a much larger influx of new mutations to be selected upon. 
At higher migration rates, greater mixing allows the initial lakes to share alleles instead of developing their own genetic basis for adaptation.

To investigate in more depth how locally adaptive alleles found in the original lakes are shared between lakes, as well as how they spread to the new lakes, we defined and tracked the distribution of ``pre-existing freshwater adapted alleles'' at the beginning of each generation. 
To be considered in this category, an allele must participate in the genetic basis of local adaptation for at least one of the original lakes. 
Concretely, these are any non-neutral mutations whose frequency is above 50\% in at least one original lake and below 50\% in the marine habitat. 
Figure \ref{fig:MPFAI}A shows the distribution of the number of these alleles across generations. At $M = 0.01$, each lake has a private set of about 10 mutations nearly fixed in that lake but not elsewhere: new lakes independently acquire new adaptive alleles rather than pre-existing ones. 
At $M = 0.1$, we again observe the original lakes adapting nearly independently from each other, but now the new lakes adapt using pre-existing alleles present in the original set of lakes. 
Concurrently, the average marine individual carries $\approx 2$ pre-existing freshwater adapted alleles, standing variation which was nearly absent at $M=0.01$. As migration rate increases past this, the total number of pre-existing freshwater adapted alleles declines (Figure \ref{fig:MPFAI}C). Interestingly, the frequency of these alleles in the ocean
stays relatively constant across the reasonable rates of gene flow.

>>>>

Threespine stickleback populations provide a striking example of local adaptation to divergent habitats 
in populations that are connected by recurrent gene flow. 
These small fish occur in marine and freshwater habitats throughout the Northern Hemisphere, and in numerous cases the smaller freshwater populations have been established ``de novo'' from marine colonists. 
Independently evolved freshwater populations show similar phenotypes that have been shown to derive largely from the same standing genetic variants.
Geographic isolation prevents direct migration between the freshwater populations, strongly suggesting that these shared locally adaptive alleles are transported through the marine population.
However it is still largely unknown how gene flow, recombination, and selection jointly impact the standing variation that might fuel this adaptation.
Here we use individual-based, spatially explicit simulations to determine the levels of gene flow that best match observed patterns of allele sharing among habitats in stickleback.
We aim to better understand how gene flow and local adaptation in large metapopulations determine the speed of adaptation and re-use of standing genetic variation. 
In our simulations we find that repeated adaptation uses a shared set of alleles that are maintained at low frequency by migration-selection balance in oceanic populations. 
This process occurs over a realistic range of intermediate levels of gene flow that match previous empirical population genomic studies in stickleback. 
Examining these simulations more deeply reveals how lower levels of gene flow leads to slow, independent adaptation to different habitats, whereas higher levels gene flow leads to significant mutation load -- but an increased probability of successful population genomic scans for locally adapted alleles. 
Surprisingly, we find that the genealogical origins of most freshwater adapted alleles can be traced back to the original generation of marine individuals that colonized the lakes, as opposed to subsequent migrants. 
These simulations provide deeper context for existing studies of stickleback evolutionary genomics, and guidance for future empirical studies in this model.
More broadly, our results support existing theory of local adaptation but extend it by more completely documenting the genealogical history of adaptive alleles in a metapopulation.

@@

The canonical model for the genetics of adaptation has long been the sequential fixation of new mutations 
\citep{smith_haigh_1974, Endler_1977, Orr_2005}.
While it has proved to be a useful baseline for understanding the genetic variation we see in species today, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures 
\citep{Lai2152, Schrider2017}. 
In particular, empirical studies have increasingly identified the need to more deeply incorporate standing genetic variation into adaptation dynamics for metapopulations inhabiting an array of habitats 
\citep{Hermisson_2008, Barrett2008}. 
Populations experiencing diverse selective pressures while still exhibiting significant gene flow often result in more complex genomic signals that are still not fully understood 
\citep{Charlesworth_1997, Charlesworth_2003, Nosil_2009, Flaxman_2013, Samuk_2017}. 
Concurrently, a growing number of empirical studies have identified instances of convergent evolution using standing genetic variation 
\citep{Schrider2017, Barrett2008, Nelson_2018, Nelson_2019, Bassham_2018}. 
However, it is still not clear how variation in evolutionary processes - such as gene flow, recombination, selection and mutation -  can promote the maintenance and re-use of standing genetic variation, particularly during colonization and adaptation to new environments
\citep{nelson2017ancient, Pritchard_2010, yeaman2011genetic, Schrider2017}. 
It is similarly unclear whether variation in these evolutionary processes can determine the genetic architecture of evolving traits via standing genetic variation. 
Theoretical work is mostly limited to one or two loci \plr{CITE}, or relies on approximations of uncertain validity.
Therefore, detailed simulation models of specific systems can 
provide an important complement to empirical studies in the lab and field,
by helping us understand precisely how standing genetic variation might fuel local adaptation,
and what genomic signals we can expect to see.

The ancestral marine form of threespine stickleback fish has given rise to millions of independently derived freshwater populations in recently de-glaciated regions around the Northern Hemisphere 
\citep{Bell_1994, Thompson_1997, Cresko_2007}. 
This model organism has provided some of the earliest data showing the heterogeneous nature of divergence across genomes 
and the much more extensive use of standing genetic variation than once thought 
\citep{schluter2009genetics, Roesti_2014, Nelson_2018, Bassham_2018, Nelson_2019, Hohenlohe_2010, Terekhanova_2014, Marques_2016}.  
While geographic isolation often prevents direct migration between freshwater populations, stickleback in them frequently evolve similar phenotypes 
\citep{Cresko_2004, Colosimo_2004}. 
The most recent evolutionary genomic studies on stickleback document that while the overall dynamic of local adaptation to marine and freshwater habitats has been occurring for millions of years 
\citep{Nelson_2018}, independent local adaptation of marine individuals to freshwater environments has been observed to take place in just tens of generations \citep{Terekhanova_2014, Lescak2015evolution, Bassham_2018}. 
For example, in 1964 the Great Alaskan Earthquake caused an uplift of many islands and coastal regions throughout the Gulf of Alaska. 
Studies of stickleback populations on uplifted Middleton Island showed that newly created freshwater ponds were invaded by the surrounding marine population of stickleback which evolved the freshwater syndrome of phenotypes in less than 50 years 
\citep{Lescak2015evolution, Bassham_2018}. 
Amazingly, the portions of the genomes of these populations that showed increased divergence from the oceanic population
mirrored those previously found to differ between ocean and freshwater populations that
that have been geographically separated for thousands of years 
\citep{Hohenlohe_2010, Bassham_2018}.

But how can evolution occur at such a rapid pace?
Waiting for new mutations to arise in each lake or pond would take much longer than the decades since the Alaskan earthquake
with any plausible target size.
Even more improbable is having divergence cluster in such similar genomic regions across independent populations. 
An alternative hypothesis
is that the majority of alleles important for freshwater adaptation are maintained in the marine individuals due to recurrent gene flow from freshwater back in to marine populations. 
\citet{schluter2009genetics} proposed a conceptual model they termed the ``transporter'' hypothesis to describe the process by which 
alleles beneficial in freshwater environments are maintained at migration-selection balance in the larger oceanic population and thus available for subsequent adaptation to new freshwater habitats.
(The alleles conferring adaptation to freshwater environments are thereby ``transported'' to and reassembled in new lakes.)
The first clear example of the global reuse of such alleles in stickleback was the gene \textit{eda} which has been shown to be an important regulator for the number of bony lateral plates 
\citep{Colosimo_2004, Cresko_2004, Colosimo_2005}. 
While the low lateral plate version of this gene arose millions of years ago, it is found in much younger freshwater ponds around the Northern Hemisphere 
\citep{O'Brian_2015}. 
More recently, genome-wide haplotype analyses have provided evidence that \textit{most} regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{nelson2017ancient}.

While the growing body of population genomic data on stickleback evolution supports the transporter hypothesis, a number of questions remain. 
What are the actual population sizes, migration rates, and fitness differentials consistent with this hypothesis? 
How many differentially selected alleles exist, how many are used at any one time, and how are they arranged within the genome? 
A curious natural history observation underlying many of these questions is the fact that some newly formed freshwater habitats, such as the ponds on Middleton Island, 
are quite small and presumably the number of initial marine migrants is few,
thus presumably representing a small portion of available variation.

Here, we use individual-based forward simulations implemented in SLiM that incorporates selection on a quantitative trait
explicitly determined additively from a realistically long genome to model the stickleback metapopulation and address these questions \citep{haller2017slim,haller2018slim3}. 
We ask how variation in amount of gene flow affects the genetic architecture of local adaptation to newly created freshwater ponds.
Because we record the entire genealogy of all alleles \citep{kelleher2018efficient}, 
we can determine the distribution and abundance of haplotypes across all marine and freshwater populations to know the timing and proportion of potentially adaptive alleles that are actually utilized in each population. 
In addition, we can document how these adaptive alleles are distributed across the genomes, and as a consequence determine how this may affect the efficacy of genome scans for between-habitat differentiation. 
We find that only a few stickleback migrants per lake per generation are sufficient to maintain standing genetic variation in the ocean, allowing new freshwater populations to adapt in only tens of generations. 
We also find that continued gene flow of marine individuals into recently established lakes is not necessary for adaptation as the majority of alleles are present in the initial marine colonizers. 
Finally, we find that this rapid, local adaptation can only occur over an intermediate range of levels of gene flow. 
Low gene flow prevents variation from being 
maintained in the ocean if it is maladaptive there, while high gene flow prevents local adaptation. 
Together these simulation findings enhance and extend the existing rich empirical work on stickleback evolutionary genomics.

@@

To explore these questions, we used SLiM \citep{haller2017slim,haller2018slim3} to implement forward-time simulation of populations of individuals with explicitly represented genomes in which selection acted upon a single continuous quantitative trait. 
The details of the model were motivated by current understanding of threespine stickleback evolutionary history and demography, and included divergent selection in the two habitats which had substantial spatial structure. 
Certain aspects of the model remain simplistic due to computational constraints;
in particular, total population sizes are much smaller than in reality,
although we may capture a good picture of local dynamics.

\paragraph{Habitat and geography}
Our simulations include two habitat types -- marine and freshwater -- defined by the nature of their selective pressures, each with 5,000 diploid individuals. 
The arrangement of these habitats, depicted in Figure~\ref{fig:Geo}, roughly models a set of freshwater habitats along a stretch of coastline. 
The marine habitat is a continuous, one-dimensional range of length 25 units, while the freshwater habitat is divided into 25 discrete subpopulations (which we call ``lakes''), each connected to the marine habitat at regularly spaced intervals (positions $i - 1/2$ for $1 \le i \le 25$).

Divergent selection is mediated by a single quantitative trait with different optima in marine and freshwater habitats. 
This situation roughly models the cumulative effect of the various phenotypes thought to be under divergent selection between the habitats, such as armor morphology, body size, craniofacial variation and opercle shape.
The optimal trait values in the marine and freshwater habitats are $+10$ and $-10$ respectively, and fitness of a fish with trait value $x_\text{ind}$ in a habitat with optimal value $x_\text{opt}$ is determined by a Gaussian kernel with standard deviation 15, i.e.,

@@

We use SLiM to simulate a Wright--Fisher population with non-overlapping generations and a fixed population size of 5,000 diploid individuals in each habitat. 
Each generation, the two parents of each new offspring are chosen proportional to their fitness 
(unlike actual stickleback, all individuals are hermaphroditic), and the contributing genomes are produced by Poisson recombination with an average of one crossover per chromosome per generation ($10^{-8}$ per locus per generation). 
Since the total population across \emph{all} 25 lakes is fixed at 5000, and the Wright--Fisher model assumes unrealistic global population regulation, we normalize the fitnesses of each individual so that approximately 200 offspring are generated in each lake, each generation. 
(A better implementation would use SLiM's ``non-Wright-Fisher'' model type,
which was not available when we wrote the simulations; but we do not expect the difference to affect results.)
To do this, we divide fitness values of each freshwater individual by the mean fitness in their lake, so that the mean fitnesses of all lakes are equal before selection happens.

@@

We used SLiM's ability to record \emph{tree sequences} \citep{haller2018treesequence,Kelleher2016} to output the genealogical history of all individuals at the time of introduction of new lakes, at the time of adaptation, and at the end of the simulation. This allowed us to directly query the true origins of adaptive alleles. 
In addition, it allowed for much larger simulations by avoiding the computationally expensive task of simulating neutral mutations which were retroactively added to the gene trees at a rate of $10^{-8}$ per locus per generation, as described in \citet{kelleher2018efficient}.
The tree sequence output by each simulation allows us to explore the origin of the genetic basis of adaptation in the new lakes. 
To do this, we constructed the genealogical tree relating all extant chromosomes at each locus along the genome. 
Using these trees we classified each adaptive allele, in each genome in the new lakes at the time of adaptation, into four categories:

@@

We used neutral mutations to calculate $F_{ST}$ for each locus, which we then averaged in windows as a measure of between-population relative differentiation. 
Concretely, if $p_f$ and $p_m$ are the frequencies of a given mutant allele in the freshwater and marine habitats, respectively, and $\bar p = (p_f + p_m)/2$, 
then we compute $F_{ST}$ for that mutation as $1 - (p_f (1-p_f) + p_m (1-p_m))/ (2 \bar p (1-\bar p))$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results \& Discussion}

@@

To observe the impact of gene flow on selection in the new lakes, we varied the ocean--lake migration rate, $m$, across separate simulations from $5 \times 10^{-5}$ to $5 \times 10^{-1}$. 
Below, we often refer to these migration rates in terms of the number of migrants per lake, per generation, which we denote $M$. Since each lake contains 200 individuals, $M=200m$. 
Many aspects of adaptation changed substantially across this range, including the speed of adaptation, degree of sharing of adaptive alleles between lakes, and the population genetic signals left behind. 
At very low rates of gene flow, each new lake's population adapted almost completely independently through \emph{de novo} mutation, which took a very long time ($\approx 20,000$ generations). 
At very high rates of gene flow, local adaptation was constrained by the large influx of locally maladaptive alleles. % \citep{bolnick2007natural}. 
Between these two extremes, genetic variation that allowed adaptation to freshwater habitats could move relatively easily between lakes. 
Perhaps surprisingly, only a few migrants per generation from the lakes to the ocean were needed to maintain sufficient genetic variation in the ocean to dramatically accelerate adaptation in new lakes.

@@

\begin{figure}
    \begin{center}
        \includegraphics{Final_Plots/Pheno_Time.pdf}
          \caption{ 
        Mean individual trait values in the marine habitat (red line),
                the original lakes (tan lines; average thicker),
                and the new lakes (black lines; average thicker),
                across the course of two simulations, with migration rates of
                \textbf{(A)} $M=0.01$ and
                \textbf{(B)} $M=0.1$ 
                migrants per lake per generation, respectively.
                Optimal trait values in the two habitats are at $\pm 10$.
                    Analogous plots for other migration rates
                    are shown in Figure \ref{fig:phenotype_time_supp}.
        }
          \label{fig:phenotype_time}
    \end{center}
\end{figure}

As shown in Figure~\ref{fig:MeanPhenotype}, local adaptation occurred in all simulations with the exception of the highest gene flow at which half of each population was composed of migrants. 
Figures \ref{fig:phenotype_time} and \ref{fig:phenotype_time_supp} show how mean trait values in freshwater and marine populations diverged over time, until the trait means were close to the optimal values in each habitat. 
Establishment of new alleles in the lakes during the first few thousand generations manifests as jumps in the mean trait value which move the trait by an amount of order 1 every few hundred generations (Fig. \ref{fig:phenotype_time}).
At the lowest rate of gene flow, $M = 0.01$, differences at around 20 commonly polymorphic sites (about 10 that shift the trait in each direction) were responsible for most of the adaptive differences between freshwater and marine habitats. 
As expected, increasing migration rate decreased differentiation between habitats: as seen in Figure~\ref{fig:Fst}, $F_{ST}$ between marine and freshwater habitats at neutral sites steadily declines as migration increases.

\begin{figure}
    \begin{center}
        \includegraphics[width=\textwidth]{Final_Plots/Fst_Genome_faa.pdf}
          \caption{
        Average $F_{ST}$ in windows of 500Mb between:
        \textbf{(top)} marine habitat and old lakes;
        \textbf{(middle)} marine habitat and new lakes; and
        \textbf{(bottom)} old and new lakes.
        Each plot shows $F_{ST}$ values for a separate simulation,
        with columns corresponding to increasing gene flow from left to right.
        $F_{ST}$ is calculated between all marine individuals, and all lakes pooled together.
        All locations of pre-existing freshwater adapted alleles have been highlighted by 
        partially transparent pink dashed vertical lines, so darker shades of pink show more 
        freshwater adapted alleles at that location.
     } \label{fig:Fst}
    \end{center}
\end{figure}

Adaptation occurred much more quickly at higher migration rates, both in the old and new sets of lakes. We measured this ``time to adaptation'' as the number of generations until average trait values in old and new lakes were within 0.5 of each other, shown in Figure~\ref{fig:TimeToAdaptation} for different rates of gene flow. Adaptation of new lakes took over 18,000 generations at the lowest rate of $M = 0.01$, while at $M = 1$ migrant per lake per generation, new lakes adapted in just under 60 generations.

@@

At low migration rates, the \emph{initial} period of adaptation takes roughly 25 times longer for lakes than it does for the ocean. 
This difference occurs because in the absence of gene flow, each lake must wait for its own novel mutations to arise in order to adapt. 
Because the marine habitat is continuous, with 25 times more individuals than any one lake, there is a much larger influx of new mutations to be selected upon. 
At higher migration rates, greater mixing allows the initial lakes to share alleles instead of developing their own genetic basis for adaptation.

To investigate in more depth how locally adaptive alleles found in the original lakes are shared between lakes, as well as how they spread to the new lakes, we defined and tracked the distribution of ``pre-existing freshwater adapted alleles'' at the beginning of each generation. 
To be considered in this category, an allele must participate in the genetic basis of local adaptation for at least one of the original lakes. 
Concretely, these are any non-neutral mutations whose frequency is above 50\% in at least one original lake and below 50\% in the marine habitat. 
(So, ``captured'' alleles as defined above are pre-existing, but not all pre-existing alleles are captured.)
Figure \ref{fig:MPFAI}A shows the distribution of the number of these alleles across generations. At $M = 0.01$, each lake has a private set of about 10 mutations nearly fixed in that lake but not elsewhere: new lakes independently acquire new adaptive alleles rather than pre-existing ones. 
At $M = 0.1$, we again observe the original lakes adapting nearly independently from each other, but now the new lakes adapt using pre-existing alleles present in the original set of lakes. 
Concurrently, the average marine individual carries $\approx 2$ pre-existing freshwater adapted alleles, standing variation which was nearly absent at $M=0.01$. As migration rate increases past this, the total number of pre-existing freshwater adapted alleles declines (Figure \ref{fig:MPFAI}C). Interestingly, the frequency of these alleles in the ocean
stays relatively constant across the reasonable rates of gene flow.

Figure~\ref{fig:MPFAI}B shows the distribution, through time, of the mean \emph{percentage} of currently-defined freshwater adapted alleles that each genome in each of the populations carries. 
If all individuals across lakes carried the same set of alleles determining their trait value, this would be 100\%. At the lowest migration rate ($M = 0.01$), each genome in the original lakes have almost exactly~$1/25^{th}$ of the total number of pre-existing freshwater adapted alleles -- this is because each one of the 25 lakes has adapted with a unique set of alleles. 
Since these are \emph{pre-existing} alleles, the value is zero for introduced lakes. 
Figure \ref{fig:MPFAI}A shows us that at $0.1$ migrants per lake per generation and above, the average individual across the new lakes has nearly the same amount of pre-existing freshwater adapted alleles as individuals across the old lakes. 
As expected, the genetic basis of the freshwater phenotype seems to simplify as migration increases -- higher rates of migration allow adaptive alleles of larger effect to travel more efficiently through the population, even though they are deleterious in the ocean.
